---
title: "Lesson 14 and 15: Python for geo-scripting"
author: "Jan Verbesselt, Jorge Mendes de Jesus, Aldo Bergsma, Eliakim Hamunyela"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
---

# Introduction

## Getting started with Python within OS-GEO live linux

- Launch osgeo live from vmware viewer, login.
- Open the terminal and type the following to start python and find out what the python version is:

To start Python:
```{r, engine='bash', eval=FALSE}
python
```

An example script to find out what the installed Python version is ([more info in question asked on stackoverflow](https://stackoverflow.com/questions/1093322/how-do-i-check-what-version-of-python-is-running-my-script))
```{r, engine='python'}
import sys
print sys.version #parentheses necessary in python 3. 
```

To exit python:
```{r, engine='bash', eval=FALSE}
exit()
```

- Open the python script within a Python editor

## Python and points

Below we will explain step by step what is done is the python script so that you
can work on the script from home.

```{r, engine='python'}
from osgeo import ogr
from osgeo import osr

# Create a point geometry
wkt = "POINT (173914.00 441864.00)"
pt = ogr.CreateGeometryFromWkt(wkt)
print(pt)
```

# A simple example

```{r, engine = 'python'}

```

