---
title: "Lesson 14 and 15: Python for geo-scripting"
author: "Jan Verbesselt, Jorge Mendes de Jesus, Aldo Bergsma, Eliakim Hamunyela"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 4
    number_sections: true
    highlight: pygments
---

# To do 

## Morning: Self-study

Reminder: Self-study is critical for the completion of the excercises at the end of the tutorial!

- Optional (!!!): if you have not followed the "Python course", it is highly recommended to follow the Python Refresher via: http://www.codecademy.com/tracks/python

- Follow and complete:

    * Thursday morning: the Python intro below and https://geoscripting-wur.github.io/PythonWeek/VectorPython.html
    * Friday morning: https://geoscripting-wur.github.io/PythonWeek/RasterPython.html

## Afternoon: Feedback and discussion

Slides for during the lesson:
https://geoscripting-wur.github.io/PythonWeek/Slides.html

# Learning outcomes

* knowing how to handle spatial data using Python
    * vector data handling
        * creating a point, writing and modifying a shape file
    * raster data handling
        * reading and writing raster data
        * calculating indices
        * projection raster data

# Intro

Using Python without ArcPy?! Why?!

- Wide user community and support
- Free
- Flexiblility
- Open-source

How?! via:
- GDAL/OGR
- GEOS

Have a look at this question on GIS StackExchange:

- https://gis.stackexchange.com/questions/34509/alternatives-to-using-arcpy
- https://gis.stackexchange.com/questions/16657/clipping-raster-with-vector-layer-using-gdal

# Getting started with Python within a Linux OS

- Launch [OS-GEO live](http://live.osgeo.org/en/index.html) from the VMware player and login.

** Question: Do you know wat OS-GEO is?

- Open the Terminal and type the following to check the installed GDAL version:

```{r, eval = FALSE, engine='bash'}
## from R: system("gdal-config --version")
## From the terminal:
gdal-config --version
```

- type the following to start python and find out what the python version is:

```{r, eval=FALSE}
python # type this in the terminal to start the python interpreter
```

An example script to find out what the installed Python version is ([more info in question asked on stackoverflow](https://stackoverflow.com/questions/1093322/how-do-i-check-what-version-of-python-is-running-my-script))

```{r, engine='python'}
import sys
print sys.version #parentheses necessary in python 3. 
```

To exit python:
```{r, eval=FALSE}
exit()
```

- Open the python script within a Python editor

# A short Python refresher

## Finding help

```{r, engine='python', eval=FALSE}
import sys
help(sys)
help(1)
```

**Question**: What does this mean `___ ___` around words: e.g: `___doc___`

Try out the following!!!

```{r, engine='python'}
help('hamster')
```

see also: 

* http://www.rafekettler.com/magicmethods.html
* https://stackoverflow.com/questions/1090620/special-magic-methods-in-python

## Finding information via Pydoc

Go to: https://docs.python.org/2/library/pydoc.html

```{r, engine='bash', eval=FALSE}
pydoc -p 1234
echo "pydoc server ready at http://localhost:1234/"
```

Then go to  `http://localhost:1234/` via your preferred browser.

## Numbers and variables

*Question*: What is the difference between 10 and 10.0 when dealing with datatypes in Python???

```{r, eval=FALSE, message=FALSE, echo=FALSE}
# 10 -> Integer -> Natural + negative numbers
## Int
# 10.0 -> Floating point -> Real number() -> "digits before and after the decimal point"
## Float
```

```{r, engine='python'}
print(int(10.6))
```

Variable is a storage location or symbolic name to a value e.g.

```{r, engine='python'}
building = 'Gaia'
buildingNumber = 101
'Gaia'
"doesn't"
'Gaia' + 'is in Wageningen'
```

There is no need to say or define the datatype, `python` has a loose type variable declaration. 

*Walk like a duck and swims like a duck and quacks like a duck I call it a duck*

Python is basically a list of objects:
List are organised with indexes. E.g. 

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine = 'python')
```

## Lists

```{r campus, engine='python'}
campus = ['Gaia','Lumen', 'Radix', 'Forum']
# how to can we print Forum?
print(campus[3])
# how to access the end of the list (while having no idea how big it is)
print(campus[-1])
# how to access the first 3 items
print(campus[0:3])
```

Appending, inserting, extending and steps:
```{r, engine='python'}
campus = ['Gaia','Lumen', 'Radix', 'Forum']
campus.append("Atlas")
campus.insert(1,"SoilMuseum")
campus.extend(["Action","Vitae", "Zodiac"])
print campus
print campus[::2]
## list[start:end:step]
```

*Question?*: What are the major differences between Append/Extend?
**Question:** What building is campus[-2]?

## Dictionaries, loops, if/else

Let there be Dictionaries... Dictionary is an unordered set of key:value pairs. Like in the dictionary, 'food':'voedsel'.

```{r, engine='python'}
# dictionary
campusDic = {101:'Gaia',
             100:'Lumen',
             107:'Radix',
             102:'Forum',
             104:'Altas'}
print campusDic[102]
```

Loops: watch out here with code intendation. Python uses Intendation to define code block and not `{}` like other languages.
Print building has to be *intended* normally 1 tab or 4 spaces (recommended).

```{r, engine='python'}
campus = ['Gaia','Lumen', 'Radix', 'Forum']
for building in campus:
    print building
```


Here, `building` is a variable that will contain any item in the campus list.

Generic loops in python have to interact over a sequence of objects e.g.

```{r, engine='python'}
range(5)
for number in range(5):
    print number
```

Object interaction and functional programming is an important part of python programming and its tools are extensive.
If/else:

```{r, engine='python'}
x = 3
if x < 3:
    print "below 3"
else:
    print "above 3"
```

```{r, engine='python'}
x = 3
if x == 1:
    print "it is one"
elif x==2:
    print "it is two"
elif x==3:
    print "it is three"
else: 
    print "above 3"
```

## Functions

A function is a section of code that does something specific that you want use multipletimes without having to type the full function again but just call the function by its name.

```{r}
def printPotato():
    print "potato"
printPotato()
```

Functions accept arguments and return variables e.g.:

```{r}
def printPotato(something):
    print something
printPotato("test")
```

`return` is used to indicate what you want to obtain from the function, and you can `return` multiple items.

```{r}
def times3(number):
    tmp = number*3
    return tmp, number
print times3(4)
```

## Importing modules

Try this!

```{r, engine='python', eval=FALSE}
import this
```

```{r, engine='python', eval=FALSE}
from __future__ import braces
```



```{r}
import math
print dir(math)
```

- The best way is to check documentation: https://docs.python.org/2/.
- Modules are Python's butter and bread. 
- A module contains code that can be used or excuted.

Basically 2 ways to load a module:
```{r}
import math
print math.pi

from math import pi
print pi
```

Why is `import math` the best way to import modules?

```{r, eval=FALSE, echo=FALSE}
from numy import pi
from math import pi
```

## Some important internal modules:

- `os`: Access to operating system features
- `os.path`: Manipulating of file names
- `sys`: System specific configuration
- `glob`: Filename pattern matching
- `math`: Mathametical functions
- `datetime`: Date/Time manipulation

*Question*: What is the difference between `os` and `os.path`?

Some examples:

```{r, eval=FALSE}
import glob
glob.glob("*")
```

```{r}
from datetime import timedelta,date
delta = timedelta(days=7)
print date.today()
print date.today()+delta
```

## File access

Very simple for 99% of the cases
Write something to file:

```{r, engine='python'}
fileObj = open('test.txt','w')
fileObj.write('some simple text')
fileObj.close()
```

```{r}
fileObj = open('test.txt','r')
a = fileObj.read()
print a
fileObj.close()
```

**Question**: What does `r` and `w` mean?

## Error handling

Sometime problems occur... Errors detected during execution are called *exceptions*

Good code deals with exceptions:

```{r, eval=FALSE}
>>> >>> open("/foo0")
Traceback (most recent call last):
  File "<stdin>", line 1, in <module>
IOError: [Errno 2] No such file or directory: '/foo0'
```

The file doesn't exist, so the script stops and outputs an ugly msg

How to deal with this I/O error??? Good progamming!!!

```{r, engine='python'}

try:
    open("foo")
except IOError:
    print "no file"

## we can be more precise:
try:
    open("/foo")
except IOError:
    print "no file"
```

## Python statements

- `if`
- `for`
- `while`
- `try`
- `class` which executes a block of code and attaches its local names to a `class`, for use in object oriented programming
- `def` which defines a function or statement
- `with` which encloses a code block within a context manager
- `pass` statement, which serves as a `NOP` (no operation)
- `assert`, used during debugging to check for conditions that ought to apply
- `yield`
- `import`

# Handling Vector data with Python

Go to [Vector Data handling in Python](https://geoscripting-wur.github.io/PythonWeek/VectorPython.html)

# Raster data with Python

Go to [Raster Data handling in Python](https://geoscripting-wur.github.io/PythonWeek/RasterPython.html)

# R from Python

A very simple example using the Python Rpy2 module (only possible on Mac/Linux):
```{r, engine='python', eval =FALSE}
import rpy2.robjects as robjects
pi = robjects.r['pi']
print(pi[0])
```

# More info
- A great book: [Python Geospatial Development](http://www.amazon.com/Python-Geospatial-Development-Second-Edition/dp/178216152X)
- [The official python tutorial](https://docs.python.org/2/tutorial/)
- [Python Tutorial](http://www.tutorialspoint.com/python/)
- [Stack OverFlow Python](https://stackoverflow.com/questions/tagged/python)
- [GIS Stack Exchange](https://gis.stackexchange.com/)

