---
title: "Raster data handling with Python"
author: "Jan Verbesselt, Jorge Mendes de Jesus, Aldo Bergsma"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: united
    toc: true
    toc_depth: 4
    number_sections: true
    highlight: pygments
---

# Handling Raster data with Python

## Intro

Have a look at this question on GIS StackExchange:

https://gis.stackexchange.com/questions/34509/alternatives-to-using-arcpy

## Dataset

The data set for the raster python session can be downloaded from:
https://www.dropbox.com/s/rsc4lzkd3t2adq5/ospy_data5.zip?dl=0

Data is also available from http://www.gis.usu.edu/~chrisg/python/2009/lectures/

## Modules and drivers for raster data

`GDAL`, the wonder for raster and! vector data (via `OGR`) handling

## Reading, analysing, and writing raster data

First, the `GDAL` module is imported together with gdalconst that contains serveral default values.
Below we just import two but more can be imported by `from osgeo.gdalconst import *`

Let's get started and see comments here within the script and the output at the end of the script:

```{r, engine='python', en}
## we define a helper function do find out what the line number is in the code
## it is not necessary to type in the following 4 lines below:
import inspect
def lineno():
    """Returns the current line number in our program."""
    return inspect.currentframe().f_back.f_lineno

## The script starts here:
import os  # for finding the working directory
from osgeo import gdal
from osgeo.gdalconst import GA_ReadOnly, GDT_Float32
# First steps is to register the drivers
print "the GDAL version"
print(gdal.__version__)
gdal.AllRegister()

# get working directory to double check in which directory we are working
os.getcwd()

# Opening Erdas Imagine Images (*.img)
driver = gdal.GetDriverByName('HFA')
filename = 'data/ospy_data5/aster.img'
dataSource = gdal.Open(filename, GA_ReadOnly)
# GDAL also works with file formats like OGR
# Question: Where does GA_ReadOnly comes from?
# A. Is there a GA_Write?! Answer?
# B. ow can we find out what the options are of the object "dataSource?""

# More info about the dataSource:
dataSource.RasterXSize
dataSource.RasterYSize
dataSource.RasterCount
dataSource.GetProjection()

# Printed Nicely:
print "information about aster.img"
print 'Driver: ', dataSource.GetDriver().ShortName,'/', \
      dataSource.GetDriver().LongName
print 'Size is ',dataSource.RasterXSize,'x',dataSource.RasterYSize, \
      'x',dataSource.RasterCount
print 'Projection is ', dataSource.GetProjection()
geotransform = dataSource.GetGeoTransform()
if not geotransform is None:
    print 'Origin = (',geotransform[0], ',',geotransform[3],')'
    print 'Pixel Size = (',geotransform[1], ',',geotransform[5],')'

## dataSource has several important properties (use dir to have a clue)
## type the following in the terminal and run it line by line:
dir(dataSource)
type(dataSource)
str(dataSource)

band1 = dataSource.GetRasterBand(1)
# dir(band1)
# help(band1.ReadAsArray)
# help(gdal.Band.ReadAsArray)

band1Arr = band1.ReadAsArray(0,0,dataSource.RasterXSize, dataSource.RasterYSize)
## 0,0 is the offset

# Question: Why are the () missing from RasterXSize and from RasterYSize?
# In object oriented programming, objects have methods (functions) and attributes
# Give an example of an attribute and of a method?

## so what is band1Arr???
print "type band1Arr:"
print type(band1Arr)

## It is a Numpy array with all the numpy array functionalities

band1Arr.min()
band1Arr.max()
band1Arr.mean()

# help(band1Arr.view)
print band1Arr.view()

# Let's get all bands
band2Arr = dataSource.GetRasterBand(2).ReadAsArray(0,0,dataSource.RasterXSize, dataSource.RasterYSize)
band3Arr = dataSource.GetRasterBand(3).ReadAsArray(0,0,dataSource.RasterXSize, dataSource.RasterYSize)

## Comment:
## We are reading 1 band per array. We could have used the ReadAsArray at the dataSource level, generating a multidimensional array
## How?

# First error message (see below)
print "This is linenumber", lineno()
ndvi = (band3Arr-band2Arr)/(band3Arr+band2Arr)

#Question: What is the result of the following division???
#1/2
#1.0/2.0

import numpy as np
band1Arr=band1Arr.astype(np.float32)
band2Arr=band2Arr.astype(np.float32)
band3Arr=band3Arr.astype(np.float32)

# What is "np" in this specific numpy import?

# How do we avoid division by zero?
# Solution:
mask = np.greater(band3Arr+band2Arr,0)
ndvi = np.choose(mask,(-99,(band3Arr-band2Arr)/(band3Arr+band2Arr)))
print "NDVI min and max values", ndvi.min(), ndvi.max()

# How do we get the real minimum value?
ndvi[ndvi>-99].min()

# Output an image
outDataSet=driver.Create('data/ndvi.img', dataSource.RasterXSize, dataSource.RasterYSize, 1, GDT_Float32)
outBand = outDataSet.GetRasterBand(1)
outBand.WriteArray(ndvi,0,0)
outBand.SetNoDataValue(-99)
## Questoin where do we get the image projection info from???
outDataSet.SetProjection(dataSource.GetProjection())
## Finally let's save it... or like in the OGR example flush it
outBand.FlushCache()
outDataSet.FlushCache()
# or
# outDataSet = None
# it is not easy to know to find out what the options 
# are linked to a certain method approach
```

**Question**: How can we write a Geotif file?

An interesting www to check out if something unexpected occurs:
http://trac.osgeo.org/gdal/wiki/PythonGotchas


We can quickly check what we have done with R:
```{r, engine='R', message=FALSE}
library(raster)
b <- raster('data/ndvi.img')
a <- brick('data/ospy_data5/aster.img')
projection(b)
hist(b, 1, maxpixels=1000, plot=TRUE)
plotRGB(a, 3, 2, 1, stretch='hist')
```

## Using QGIS and Python

Here is a nice clear example how to get started:
- http://www.qgistutorials.com/nl/docs/getting_started_with_pyqgis.html

- [QGIS and Python tutorial](http://www.qgisworkshop.org/html/workshop/python_in_qgis_tutorial2.html)

# More Information
- https://pcjericks.github.io/py-gdalogr-cookbook/raster_layers.html
- Great basic tutorial: http://www.gdal.org/gdal_tutorial.html
- [Handling raster data with GDAL](http://www.gis.usu.edu/~chrisg/python/2009/)
- [GDAL Python info and bindings](http://gdal.org/python/osgeo.gdal_array-module.html#BandReadAsArray)
- yet another example https://borealperspectives.wordpress.com/2014/01/16/data-type-mapping-when-using-pythongdal-to-write-numpy-arrays-to-geotiff/

# Info about Python editors:
- https://stackoverflow.com/questions/8305809/is-there-something-like-rstudio-for-python (check out `Spyder` or `iPython`)
