---
title: "[WUR GRS-51806 Geoscripting](https://geoscripting-wur.github.io/)"
subtitle: "Week 3: Raster data handling with Python"
author: "Jan Verbesselt, Jorge Mendes de Jesus, Aldo Bergsma"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: yes
    css: custom.css
    highlight: tango
    toc: yes
    toc_depth: 2
---

# Handling Raster data with Python

## Intro

Key modules in Python for Raster data handling:

* Developped for high-level spatial analysis:
`PySAL`: https://pypi.python.org/pypi/PySAL

Have a look at this question on GIS StackExchange:
https://gis.stackexchange.com/questions/34509/alternatives-to-using-arcpy

## Dataset

The data set for the raster python session can be downloaded from:
https://www.dropbox.com/s/rsc4lzkd3t2adq5/ospy_data5.zip?dl=0

Data is also available from http://www.gis.usu.edu/~chrisg/python/2009/lectures/

## Modules and drivers for raster data

`GDAL`, the wonder for raster and! vector data (via `OGR`) handling

## Reading, analysing, and writing raster data

See here for a full [reproducible Python outline](http://nbviewer.ipython.org/github/GeoScripting-WUR/PythonWeek/blob/gh-pages/Raster%20data%20handling.ipynb).

An interesting www to check out if something unexpected occurs:
http://trac.osgeo.org/gdal/wiki/PythonGotchas

### Via Python

```{r, engine = 'python', eval=TRUE}
# import modules
from osgeo import gdal
from osgeo.gdalconst import GA_ReadOnly, GDT_Float32
import numpy as np

# open file and print info about the file
# the ¨../¨ refers to the parent directory of my working directory
filename = 'data/ospy_data5/aster.img'
dataSource = gdal.Open(filename, GA_ReadOnly)

print "\nInformation about " + filename 
print "Driver: ", dataSource.GetDriver().ShortName,"/", \
      dataSource.GetDriver().LongName

```

### Via R

We can quickly check what we have done with R:
```{r, engine='R', message=FALSE, eval=FALSE}
library(raster)
a <- brick('data/ospy_data5/aster.img')
plotRGB(a, 3, 2, 1, stretch='hist')
```

```{r, eval = FALSE}
b <- raster('data/ndvi.tif')
extent(b)
projection(b)
hist(b, 1, maxpixels=500, plot=TRUE)
```

## Using QGIS and Python

Here is a nice clear example how to get started:

* http://www.qgistutorials.com/nl/docs/getting_started_with_pyqgis.html
* [QGIS and Python tutorial](http://www.qgisworkshop.org/html/workshop/python_in_qgis_tutorial2.html)

# Assignment

For this scripting challenge I have downloaded a Landsat image with all bands process to surface reflectance (Level 1T). You can download it from [here](https://www.dropbox.com/s/zb7nrla6fqi1mq4/LC81980242014260-SC20150123044700.tar.gz?dl=0). Unzip the file and you will see that it contains all the invidual bands. 

Now, write a well-structured and documented script where you define a function to derive the NDWI and derive it from the landsat image and reproject the image to Lat/Long WGS84 (`x.ImportFromEPSG(4326)`).

`NDWI = band 4 - band 5 / band 4 + band 5` ([more info about](http://nsidc.org/data/docs/daac/nsidc0332_smex03_srs_ndvi_ndwi_ok.gd.html) `NDWI`)

A clean and well structured script is critical here.

## Deadline

Upload your documented and well structured Python script to a GitHub repository and send the link before 1030 on Monday 26/01.

## Landsat data

You can easily order landsat data via:

  * https://earthexplorer.usgs.gov/
  
Once the have the Landsat ID you can order further processed data (e.g. resampled, or as geotif, etc.) via:

  * http://espa.cr.usgs.gov/index/

# More Information
- http://complex.luxbulb.org/howto/spatial-analysis-python
- [Blog about Python for GeoSpatial data analysis](http://www.digital-geography.com/python-for-geospatial-data-analysis-part-ii/?subscribe=invalid_email#477)
- https://pcjericks.github.io/py-gdalogr-cookbook/raster_layers.html
- Great basic tutorial: http://www.gdal.org/gdal_tutorial.html
- [Handling raster data with GDAL](http://www.gis.usu.edu/~chrisg/python/2009/)
- [GDAL Python info and bindings](http://gdal.org/python/osgeo.gdal_array-module.html#BandReadAsArray)
- yet another example https://borealperspectives.wordpress.com/2014/01/16/data-type-mapping-when-using-pythongdal-to-write-numpy-arrays-to-geotiff/

# Info about Python editors:
- https://stackoverflow.com/questions/8305809/is-there-something-like-rstudio-for-python (check out `Spyder` or `iPython`)
