---
title: "Reprojecting a raster from Lat/long to UTM"
author: "Eliakim Hamunyela and Jan Verbesselt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    theme: united
    toc_depth: 4
    number_sections: true
---

# Introduction

OSGEO Python approach - to clip a raster (lat/long) into UTM zones and reproject individual clipped raster to their UTM zone. 

Once they are all in their UTM zone, they can be joined again (mozaic).

# Python

## Check your Python version

```{r, engine='python'}
import sys ## checking the Python version
print sys.version #parentheses necessary in python 3. 
```

## Read in shape file and convert ESRI projection Wkt format to PROJ.4
Doing the same via `Python`:

```{r, engine='python'}
import os
from osgeo import ogr
import osgeo.osr
srs = osgeo.osr.SpatialReference()

Llutm = r"TiffRastersAndVectorShapeData/VectorData/psUTMGridZonesWorld.shp"

## open shape file
driver = ogr.GetDriverByName('ESRI Shapefile')
dataSource = driver.Open(Llutm, 0) # 0 means read-only. 1 means writeable.

# Check to see if shapefile is found.
if dataSource is None:
    print 'Could not open %s' % (Llutm)
else:
    print 'Opened %s' % (Llutm)
    layer = dataSource.GetLayer()
    featureCount = layer.GetFeatureCount()
    print "Number of features in %s: %d" % (os.path.basename(Llutm),featureCount)

## extract the info of the UTMZonePro Layer and feed into the section below
dataSource = ogr.Open(Llutm)
daLayer = dataSource.GetLayer(0)
layerDefinition = daLayer.GetLayerDefn()
for i in range(layerDefinition.GetFieldCount()):
    print layerDefinition.GetFieldDefn(i).GetName()

print "print a specific layer"
print layerDefinition.GetFieldDefn(6).GetName()

## Feature
feature = daLayer.GetFeature(0)
id = feature.GetField('UTMZonePro')
print("print id")
print(id)
#import from wkt
srs.ImportFromWkt(id)
 
#create proj4
out = srs.ExportToProj4()
print(out) 
```

**Here the PROJ.4 file is not complete**

## Reprojecting - In progress

```{r, engine='python'}
import osgeo.osr
srs = osgeo.osr.SpatialReference()
 
#create coordinate string. 
#The wkt from our example is not complete, so we use this wkt from here: http://www.gdal.org/osr_tutorial.html

test  = 'PROJCS["UTM 18 (WGS84) in northern hemisphere.",GEOGCS["WGS 84",DATUM["WGS_1984",SPHEROID["WGS 84",6378137,298.257223563,AUTHORITY["EPSG",7031]],TOWGS84[0,0,0,0,0,0,0],AUTHORITY["EPSG",6327]],PRIMEM["Greenwich",0,AUTHORITY["EPSG",8902]],UNIT["DMSH",0.0174532925199433,AUTHORITY["EPSG",9109]],AXIS["Lat",NORTH],AXIS["Long",EAST],AUTHORITY["EPSG",4327]],PROJECTION["Transverse_Mercator"],PARAMETER["latitude_of_origin",0],PARAMETER["central_meridian",-84],PARAMETER["scale_factor",0.9996],PARAMETER["false_easting",500000],PARAMETER["false_northing",0]]'
print(test) 
#import from wkt
srs.ImportFromWkt (test)
 
#create proj4
out = srs.ExportToProj4()
print(out) 
```


# R

Reading in the shape file using `R`
```{r, eval=FALSE}
require(rgdal)
dsn <- "TiffRastersAndVectorShapeData/VectorData/"
ogrInfo(dsn, "psUTMGridZonesWorld")
# read in shapefiles
latlongutmgrid <- readOGR(dsn, "psUTMGridZonesWorld")
# note that readOGR will read the .prj file if it exists
print(proj4string(latlongutmgrid))
# info about UTM zones
latlongutmgrid$UTMZonePro[1]
out <- latlongutmgrid$UTMZonePro[1]

## export data for this zone and then read it in via Python
write.table(out, "TiffRastersAndVectorShapeData/VectorData/ZoneInfo.csv", 
            col.names = FALSE, row.names = FALSE, quote = FALSE)
```

# Summary - What have we learned

# Acknowledgements and more info

- http://www.gdal.org/osr_tutorial.html

- http://spatialnotes.blogspot.nl/2010/11/converting-wkt-projection-info-to-proj4.html

- https://pcjericks.github.io/py-gdalogr-cookbook/


<!---
### comments not processed by the kml file
I've had many occasions where I've had WKT-style projection text (e.g., the contents of the *.prj files that normally accompany ESRI Shapefiles, or the srtext field in the spatial_ref_sys table in PostGIS), and I needed to get that to an equivalent Proj4-compatible projection string (and vice versa).  I've known for some time there had to be an automated way to do this...as the GDAL/OGR libraries have been doing this internally for some time.  I ran into the issue again today, so I decided to poke around a bit, and found the API functions that are available to do this...I wrote my first ever Python scripts that use the GDAL/OGR python bindings (specifically, OSR).  These scripts simply convert WKT projection text to Proj4 (wkt2proj.py) and back (proj2wkt.py).  Maybe there are some better examples elsewhere...but I didn't quite find what I was looking for.  Since my work is mostly in PHP right now, I'd love to see the GDAL/OGR PHP bindings maintained someday, but this will do for now:
-->