---
title: "Vector data handling with Python"
author: "Jan Verbesselt, Jorge Mendes de Jesus, Aldo Bergsma, Eliakim Hamunyela"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 4
    number_sections: true
    highlight: pygments
---

# Handling Vector data with Python


## Points

```{r, engine='python'}
from osgeo import ogr
# Create a point geometry
wkt = "POINT (173914.00 441864.00)"
pt = ogr.CreateGeometryFromWkt(wkt)
print(pt)
```

## Define a spatial reference

See http://spatialreference.org/ref/epsg/wgs-84/ 

```{r, engine='python'}
from osgeo import osr
##  spatial reference
spatialRef = osr.SpatialReference()
spatialRef.ImportFromEPSG(4326)  # from EPSG - Lat/long
```

## Reproject a point

```{r, engine = 'python'}
from osgeo import ogr
from osgeo import osr

## lat/long definition
source = osr.SpatialReference()
source.ImportFromEPSG(4326)

# http://spatialreference.org/ref/sr-org/6781/
# http://spatialreference.org/ref/epsg/28992/
target = osr.SpatialReference()
target.ImportFromEPSG(28992)

transform = osr.CoordinateTransformation(source, target)
point = ogr.CreateGeometryFromWkt("POINT (5.6660 -51.9872)")
point.Transform(transform)
print point.ExportToWkt()
```

## Assignment

Upload your Python script to Github (as a .py file) uploaded within the `Python` folder part of your repository for the Python week:
Create a script that:

- defines two points in Google Earth or use the two points from the R vector excercise.
- set the projection (Lat/Long, should you know now the EPSG) and reproject to the Dutch projection (See definition in the Vector Lesson)
- draw a line between the points
- create a buffer around the points of 100m
- measure the distance between these points (i.e. the distance of that line) and print it out.
- export the two point to shape file (that can be imported in QGIS)

